---
title: 'Lab 3: Draft Report'
author: "Priscilla Burity, Oscar Linares, Alissa Stover"
date: "3/16/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# An Introduction

Research Question: Which policies are most promising in reducing the crime rate: economic (e.g., increasing wages) or classic criminal deterrants (e.g., increasing arrests) (or both together)? What is the influence of contextual factors (e.g., characteristics of a population) on these relationships? 

We will be modeling our dependent variable (crime rate, `crmrte`), using the following variables as candidate independent variables. We focus on variables that represent different economic and criminal justice policies, which can be influenced via various policy decisions taken by our candidate. 

We also include a range of contextual variables that help to control for other variables that could influence the efficacy of these policies. We aim to produce a model robust to these, that can help guide decisions at the state level and help reduce crime across these counties. 

## Actionable Variables

### Economic Policy

These variables describe things we could affect with economic policy. For example, if increased wages are found to be related significantly to crime, our candiate could consider strategies such as raising the minimum wage in an attempt to lower crime. These variables speak to wages for different sectors.

- `wcon`:	weekly wage, construction
- `wtuc`:	wkly wge, trns, util, commun
- `wtrd`:	wkly wge, whlesle, retail trade
- `wfir`:	wkly wge, fin, ins, real est
- `wser`:	wkly wge, service industry
- `wmfg`:	wkly wge, manufacturing
- `wsta`:	wkly wge, state employees
- `wloc`:	wkly wge, local gov emps

The following federal wage variable will be grouped with other wages in much of our analyses for easier comparison in terms of data quality, however it is not related to minimum wage policy but rather to cost of living. Federal employee wages are adjusted by cost of living in a locality (https://www.opm.gov/policy-data-oversight/pay-leave/salaries-wages/fact-sheets/); thus, this variable could also be used as a marker of cost of living in a county. Cost of living is related to overall wealth in an area and its overall econimic development. If we find a relationship between federal wages and crime, one could predict that changing the economic development of a region may affect crime. However, this variable is confounded by the types of federal workers in an area, since federal wages are determined by a grade system that is tied to the nature of the position and time in position. A specific effort that could affect both of these - bringing more economic development and the right types of high-earning federal employees - would be to attract federal jobs to the area that have higher wages. 

- `wfed`:	wkly wge, fed employees

The following variable represents tax policy, but it could also reflect overall wealth in a county. Holding tax rate constant, counties with more wealth will have more tax revenue and vice versa. Holding wealth constant, changing the tax rate would change the tax revenue per capita. A better measure of tax policy would be the actual tax rates in each county. Since the impact of policy on this variable is confounded by the overal wealth in a county, the causal relationship would be difficult to interpret. Plus, many taxes are set at the state and federal level; the level of variation at the county-level may be limited for a small state like North Carolina. However, similar to `wfed`, this variable would be related to overall economic development. If we found a relationship between tax revenue and crime rate, it would make sense to try to change development efforts in order to affect crime.  

- `taxpc`:	tax revenue per capita 

### Criminal Justice Policy

These variables describe things we could affect with policy around the criminal justice system. Many of these policies are used as deterrants of crime, with the assumption being that people inherently understand that these punishments exist and as such will be less likely to commit crimes (in order to avoid the punishment). 

Please note that the 3 "efficiency" variables below were initially called "probabilities". However, this is a misnomer as these variables are truly ratios (e.g., the ratio of arrests to crimes reported). We interpret these as measures of the efficiency of the criminal justice system in a county. Ideally, we would want to strike a balance between arresting enough people to prevent criminal activity, and not arresting so many as to abuse the populace.

*Efficiency of arrest* 

If we find that the efficiency of arrest is associated with crime rate (with higher punishment associted with higher crime), we may consider stricter policies around police practice that encourage more arrests. 

- `prbarr`: number of convictions for every crime reported

*Efficiency of conviction* 

If we find that the efficiency of conviction is associated with crime rate (with higher conviction rates associted with higher crime), we may consider stricter policies around court practices so that more arrests lead to convictions.

- `prbconv`: number of convictions for every arrest made

*Efficiency of sentencing* 

If we find that the efficiency of sentencing is associated with crime rate (with higher sentencing rates associted with higher crime), we may consider stricter policies around judicial practices so that more convictions lead to sentencing. However, whether or not a given conviction carries a prison sentence depends on the severity of the crime. Thus, this variable may be more limited in its explanatory power than arrests and convictions. 

- `prbpris`: number of prison sentences for every conviction

We can combine the three above variables into one variable (`acj`) that is a measure of the efficiency of the criminal justice process in each county: 

acj <- `prbarr` * `prbconv` * `prbpris` 

*Severity of punishment* 

If we find that the severity of punishment is associated with crime rate (with greater severity associted with lower crime), we may consider stricter policies around judicial practices so that people receive more severe punishments if sentenced with a crime.

- `avgsen`:	avg. sentence, days 

*Number of policy officers per capita*

The number of police police officers per capita (measured by the variable `polpc`) is a very intuitive example of crime deterrant, as it in theory increases the likelyhood of being caught in the act of the crime and is also a measure of the capacity of the State to enforce the law.

As our job is to advise a policy maker, it’s very important that we feel safe about the causal interpretation of our findings. After all, the idea is to propose policies with a significant impact on people’s lives. For statistical rigor, we also need to make sure that causal effects go one way (that we have reason to believe that our independent variables have causal effects on our dependent variables, and not the reverse). A cross section of counties does not allow us to link police presence to crime rates with a causal interpretation. This is because one can expect the counties with larger crime rates in a year *t* would consider necessary to have a larger number of police officers in on the streets in year *t*, which can imply a positive correlation between presence of police and crime rates with no causal meaning. An increase in crime could result in an increase in the number of police officers *and* an increase in police officers could result in a decrease in crime. 

Thus, in this exercise we opted not to include police per capita as a candidate policy variable or covariate. 

## Contextual Variables

Since we are aiming to produce policy recommendations, the following variables may be helpful in our analysis but are not necessarily actionable. They describe county characteristics that provide important information about which policies could work across different county contexts.

### Types of Crime 

- `mix`:	offense mix: face-to-face/other

### Demographics 

The following variables describe different aspects of the people that reside in these counties.

*Urban/rural dwellers* 

- `density`:	people per sq. mile 

*Minority status*

- `pctmin80`:	perc. minority, 1980 

*Gender & Age* (young males are more likely to enter the criminal justice system)

- `pctymle`:	percent young male 


### Geography

The following variables identify which region of the state each county is in, with the assumption that counties cluster geographically in terms of their culture and other characteristics we are not explicitly measuring. Western North Carolina is along the Appalachian mountains and seems qualitatively different from much of the rest of the state. For example, it has much fewer universities compared to the rest of the state (you can see this through a quick Google Maps search). Thus, we will lean towards using `west` over the other geographical variables (especially since `urban` could be highly related to `density`). 

- `west`:	=1 if in western N.C.
- `central`:	=1 if in central N.C.
- `urban`:	=1 if in SMSA

# A Model Building Process

## Data Cleaning

### Environment

#### Import Libraries 

```{r, message=FALSE}
library(tidyverse) # for data import, manipulation, viz
library(corrplot) # for correlation matix
library(stargazer) # visualize model fit
library(skimr) # generate summary statistics
library(car) # statistics 
library(lmtest) # linear modeling
library(olsrr) # evaluating OLS regression 
```

#### Import Data

We are importing a CSV file into R. The argument `n_max` is set to the number of rows of data we are provided (91). 

```{r, message=FALSE}
data <- read_csv("crime_v2.csv", n_max = 91)
```

#### Exploratory Data Analysis: File-level data checks

Let's take a visual check of our data. We have 91 observations and 25 columns. This is what we expected based on our original csv file. 

```{r}
head(data)
dim(data)
```

All 25 columns are numeric. 

```{r}
str(data)
```

##### The Where & When of these data 

It looks like we usually have 1 observation for each of the counties -- except for county #193. We do not have any missing values in the `county` column. 

```{r}
length(unique(data$county))
table(data$county, useNA = "always")
```

Having 2 observations for county #193 is surprising. We can see that these rows appear to be exact duplicates. 

```{r}
data %>% filter(county == 193)
```

We can see that `year` appears to be constant (which is expected). We also don't have any missing values for this column. We can drop this column in future data processing steps. 

```{r}
table(data$year, useNA = "always")
```

Based on the explorations above, we create a new dataframe that has dropped our superfluous column (`year`) and the superfluous row (duplicate of county 193). 

```{r}
data2 <- data %>% distinct() %>% select(-year)
```

Now we have 90 rows and 24 variables, with one observation per county.

```{r}
length(unique(data2$county)) == length(data2$county) # check 1 obs per county
str(data2) # examine dataframe 
```

##### Missingness & Data Validity

We can see from the following table that we don't have any missing data, which is good! Please note that this is with our additional crime policy variable (`acj` = `prbarr` * `prbconv`* `prbpris`).

However, we can see by the max for each column (p100), that we appear to have a few values that are extremely high. 

For example, one county has an average sentence length (`avgsen`) of 20.7 years.

```{r}
# create new acj variable
data2 <- data2 %>% mutate(acj = prbarr * prbconv * prbpris)
```

```{r}
data2_skim <- skim(data2)
data2_skim
```

##### Preserve Clean Data

Save our clean dataset

```{r}
save(data2, file = "data_clean.rda")
```

## Exploratory Data Analysis: Variable-level data checks

### Dependent Variable: Crime Rate

Barring extreme issues with `crmrte` (crime rate), we would recommend focusing on this as our target variable. Although arrest, conviction, and sentencing rates are important, what matters most to people is the level of crime around them every day. The justice system is used to enact justice and to deter crimes. Justice is important, and making sure that those that commit crimes are punished is one way of going about achieving a just society. However, justice is a moral issue and can be seen from many different lenses. On the other hand, most people would agree that not having a lot of crime around them - and the sense of safety that comes from it - is very important. Thus we will use these quantitative methods to evaluate overall crime, rather than try to monitor the fuzzier concept of "justice". While doing so, we will also make sure to evaluate the arrest/conviction/sentencing variables to see whether they are related to different levels of crime - whether they could be having any effect as crime deterrents would be important for policymakers to know. 

We don't have enough data points to make a strong claim about the distribution of `crmtre`, however it appears to be approximately normal (with a slight skew which could be an artifact of not having any values below 0). 

```{r}
ggplot(data2, aes(crmrte)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  xlab("Crime Rate") +
  ylab("Count") +
  ggtitle("Histogram of Candidate Dependent Variable: Crime Rate")
```

From our boxplot we can see that most values are between 0.02 and 0.04, however we do have a range of values including some that are close to 0.10. We don't appear to have any spurious values. 

```{r}
ggplot(data2, aes(y = crmrte)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Crime Rate") +
  ggtitle("Boxplot of Candidate Dependent Variable: Crime Rate")
```

Our preferred variable (`crmrte`) appears to be of sufficient quality for our analyses and thus we will use it as our dependent variable. 

### Independent Variables 

#### Actionable Variables

##### Economic Policy

One of the things one can affect with economic policy are wages. In North Carolina in 1987, the minimum hourly wage was \$3.35. Given that most people work 40 hours per week, the minimum hourly wage would be \$134 per week.

In the plots below, we plot weekly wage by sector with a red vertical line indicating this minimum weekly wage.

We can see that `wrtd` (Wholesale/Retail/Trade) wages are closest to the minimum wage - which would suggest that wages in this sector would be most influenced by changes to the minimum wage. The next sector that would be influenced is `wser` (Service Industry). However, `wser` appears to have an outlier (someone is earning over $2000 weekly, which is well above the weekly wages for any other sector as well as `wser`). Additionally, `wser` may be a messier measure of income because many in that industry work in restaurants and make additional income through tips - however, some people may not report this income. Meanwhile, those working in Wholesale/Retail/Trade may be more likely to work in places like malls, which would not provide additional income as tips and would be a better measure of low-income wages. 

```{r}
# format data for subplots 
data_ep_long <- data2 %>% 
  select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc) %>%
  gather(key = var, value = value)

data_ep_long$var <- factor(data_ep_long$var,
                           levels = c("wcon", "wtuc", "wtrd", "wfir",
                                      "wser", "wmfg", "wfed", "wsta", "wloc"),
                           labels = c("Construction", 
                                      "Utilities/Transport", 
                                      "Wholesale/Retail/Trade",
                                      "Finance/Insurance/Real Estate",
                                      "Service Industry",
                                      "Manufacturing",
                                      "Federal Emp",
                                      "State Emp",
                                      "Local/State Emp"))
```

```{r}
# plot hist subplots
ggplot(data_ep_long, aes(value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~var) +
  theme_minimal() +
  ggtitle("Histogram of Economic Policy Variables: Wages by Sector") +
  geom_vline(aes(xintercept = 134, color = "red")) +
  theme(legend.position = "none")

```

```{r}
# plot boxplot subplots 
ggplot(data_ep_long, aes(x = var, y = value)) +
  geom_boxplot() +
  coord_flip() + 
  theme_minimal() +
  xlab("") + 
  ylab("Weekly Wage") +
  ggtitle("Boxplot of Economic Policy Variables: Wages by Sector")+
  geom_hline(aes(yintercept = 134, color = "red")) +
  theme(legend.position = "none")

```


One can see from the plot below that `wser` for county #185 appears to be an incorrect value - all of the other sector wages are close to the mean for all counties, except for this one (which is about 10x the mean). 

```{r}
data_wages <- data2 %>% select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc)%>% colMeans() %>% as_tibble() %>% mutate(names = c("wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc"), Group = "Mean")

wages_185 <- data2 %>%filter(wser == max(wser)) %>% select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc) %>% colMeans() %>% as_tibble() %>% mutate(names = c("wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc"), Group = "County #185")

wages <- bind_rows(data_wages, wages_185)

ggplot(data = wages, aes(x = names, y = value, color = Group)) + 
  geom_point() +
  xlab('Wage Category') +
  ylab('Wage') +
  theme_minimal() +
  ggtitle("Comparing Wages by Sector between County 185 and Other Counties")

```

##### Criminal Justice Policy

As noted above, we combined our 3 core criminal justice policy variables (`prbarr`, `prbconv`, `prbpris`) into one measure of the assertiveness of the criminal justice system. It is a measure of the ratio of arrests, convictions, and prison sentences to crimes. The closer this value is to 1, the more "efficient" one would imagine the criminal justice process to be (there is 1 arrest/conviction/prison sentece for every crime). Values lower than 1 indicate that there are crimes that go unpunished. Values higher than 1 indicate that there are extra punishments enacted for every crime. 

First, we evaluated the quality of these 3 variables separately. 

```{r}
# format data for subplots 
data_cj_long <- data2 %>% select(prbarr, prbconv, prbpris) %>%
  gather(key = var, value = value)

# convert to factor for subplot labelling 
data_cj_long$var <- factor(data_cj_long$var, labels = c("Prob Arrest", "Prob Convicted", "Prob Sentenced"))
```

We can see that we have different distributions for these, but that they all could be considered somewhat close to normal, with different levels of skew. efficiency of conviction appears to be closest to a normal distribution. 

```{r}
# plot hist subplots
ggplot(data_cj_long, aes(value)) +
  geom_histogram(bins = 50) +
  facet_grid(~var) +
  theme_minimal() +
  ggtitle("Histogram of Criminal Justice Policy Variables: efficiency Variables")
```

From our boxplots we can see that we have some extreme values for efficiency of conviction, with multiple counties having a rate above 1. As these are not true probabilities, but rather ratios, these values are not necessarily spurious. However, they imply that for each arrest, there could be multiple convictions - one county has over 2 convictions for every arrest. Only 1 county has more than 1 arrest per crime committed - most counties in fact have a much lower rate (and only put someone under arrest for about one-quarter of crimes). When it comes to setencing, counties show a much lower spread and cluster around 1 conviction resulting in a prison sentence for every 1 conviction that does not. 

```{r}
# plot boxplot subplots 
ggplot(data_cj_long, aes(x = var, y = value)) +
  geom_boxplot() +
  theme_minimal() +
  xlab("") +
  ylab("Ratio") + 
  ggtitle("Boxplot of Criminal Justice Policy Variables: Efficiency Variables")
```

Which county had the highest conviction rate? This is a county in central North Carolina. It has the highest percentage of minorities. It also appears to have the maximum wages in the service industry ($2177.068 compared to a median of \$253.12). 

```{r}
data2 %>% filter(prbconv == max(prbconv))
```

The fact that the conviction rate is so high seems odd for county #185. The other measures of efficiency (for arrest and sentencing) seem close to the mean. That of conviction rate is astronimically high. 

```{r}
data_prb <- data2 %>% select(prbarr, prbconv, prbpris) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris"), Group = "Mean")

prb_185 <- data2 %>% filter(prbconv == max(prbconv)) %>% select(prbarr, prbconv, prbpris) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris"), Group = "County #185")

prbs <- bind_rows(data_prb, prb_185)

ggplot(data = prbs, aes(x = names, y = value, color = Group)) + 
  geom_point() +
  xlab('Criminal Justice Efficiency Category') +
  ylab('Efficiency') +
  theme_minimal() +
  ggtitle("Comparing Criminal Justice Efficiencies between County 185 and Other Counties")

```

The analysis above suggests that the data for this county #185 may have been corrupted - since it has values for `wser` and `prbconv` at levels much higher than the mean, whichout any explanation for why these might have both been elevated. We cannot know for sure as we do not have any other data with which to triangulate these records, thus in our modeling we will compare models with and without this county; if it does not affect the models, we can leave in this outlier. If it does significantly, we may remove it since the data do not seem correct. 

```{r}
# plot hist
ggplot(data2, aes(y = acj)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Boxplot of Criminal Justice Policy Variables: Assertiveness") +
  ylab("Average Sentence Assertiveness")
```

We can see that when combined, all but one of the counties have an assertiveness measure below 0.2. One county has an exceptionally high score on this measure (above 0.8). 

The assertiveness variable is approximately normally distributed. 

```{r}
# plot hist 
ggplot(data2, aes(avgsen)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  xlab("") +
  ylab("Assertiveness") + 
  ggtitle("Boxplot of Criminal Justice Policy Variables: Assertiveness")
```

```{r}
data2 %>% filter(acj == max(acj))
```

We can see that this Western county has a very high arrest and conviction rate compared to the mean. However, the distance between each from the mean seems consistent. Its prison sentencing rate is much closer to the mean. 

```{r}
data_acj <- data2 %>% select(prbarr, prbconv, prbpris, acj) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris", "acj"), Group = "Mean")

acj_115 <- data2 %>% filter(acj == max(acj)) %>% select(prbarr, prbconv, prbpris, acj) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris", "acj"), Group = "County #115")

acjs <- bind_rows(data_acj, acj_115)

ggplot(data = acjs, aes(x = names, y = value, color = Group)) + 
  geom_point() +
  xlab('Criminal Justice Efficiency Category') +
  ylab('Efficiency') +
  theme_minimal() +
  ggtitle("Comparing Criminal Justice Efficiencies between County 115 and Other Counties")

```


We can see that average sentence length follows an approximately normal distribution. We have one value well above the rest (above 20; most cluster below 15). 

```{r}
# plot hist
ggplot(data2, aes(avgsen)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  ggtitle("Histogram of Criminal Justice Policy Variables: Severity") +
  xlab("Average Sentence Severity")
```

```{r}
# plot hist
ggplot(data2, aes(y = avgsen)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Boxplot of Criminal Justice Policy Variables: Severity") +
  ylab("Average Sentence Severity")
```

We can see that the same county with the highest value of the assertiveness is the same value with the highest average sentence - one may expect those to track together, as more extreme policies against crime would extend to all of these practices. This county is in Western North Carolina. It has the lowest percentage of minorities of any county (1.28%). Its density is in the lowest quartile of counties but its mix of face-to-face versus other crime is at the median. Although this county seems to be at an extreme, there is no reason to eliminate it from the data at this juncture. 

```{r}
data2 %>% filter(avgsen == max(avgsen))
```

#### Contextual Variables

### Types of Crime 

```{r}
# plot barplots
ggplot(data2, aes(mix)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  ggtitle("Histogram of Contextual Variables: Type of Crime") +
  xlab("offense mix: face-to-face/other")
```

##### Demographics 

*Urban/rural dwellers* 

```{r}
ggplot(data2, aes(density)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  xlab("Density") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Population Density")
```


```{r}
ggplot(data2, aes(y = density)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Density") +
  ggtitle("Boxplot of Contextual Variables: Population Density")
```

*Minority status*

```{r}
ggplot(data2, aes(pctmin80)) +
  geom_histogram(bins = 40) +
  theme_minimal() +
  xlab("% Minority") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Percentage Minority (1980)")
```


```{r}
ggplot(data2, aes(y = pctmin80)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("% Minority") +
  ggtitle("Boxplot of Contextual Variables: Percentage Minority (1980)")
```

- `pctmin80`:	perc. minority, 1980 

*Gender & Age* (young males are more likely to enter the criminal justice system)

```{r}
ggplot(data2, aes(pctymle)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  xlab("% Young Males") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Percentage Young Males")
```

```{r}
ggplot(data2, aes(y = pctymle)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("% Young Males") +
  ggtitle("Boxplot of Contextual Variables: Percentage Young Males")
```

- `pctymle`:	percent young male 

*Wealth*

```{r}
ggplot(data2, aes(taxpc)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  xlab("Tax Revenue Per Capita") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Tax Revenue Per Capita")
```

```{r}
ggplot(data2, aes(y = taxpc)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Tax Revenue Per Capita") +
  ggtitle("Boxplot of Contextual Variables: Tax Revenue Per Capita")
```

- `taxpc`:	tax revenue per capita 

##### Geography

```{r}
# format data for subplots 
data_geo_long <- data2 %>% select(west, central, urban) %>%
  gather(key = var, value = value)
```

```{r}
# plot barplots
ggplot(data_geo_long, aes(value)) +
  geom_bar() +
  facet_grid(~var) +
  theme_minimal() +
  ggtitle("Barplot of Contextual Variables: Geography")
```

- `west`:	=1 if in western N.C.
- `central`:	=1 if in central N.C.
- `urban`:	=1 if in SMSA


### Relationships between Variables

One would expect variables within each of the above groups to be correlated to each other. 

This is a correlation matrix of all of our variables, using Spearman's due to the fact that not all variables follow a normal distribution. 

Our original target variables for economic policy -- those related to wages -- appear to all be positively correlated with crime, which is opposite our original prediction. Rather than seeing that increased wages are related to reduced crime, we are seeing that higher wages is related to higher crime. This unexpected direction could be explained in part by the strong positive relationship between the wage variables and `density` and between `crmrte` and `density`. Thus, we will not include the wage variables in early models - if we include them, they will only be included along with covariates such as `density`. 

```{r}
data_corr <- data2 %>% select(-county) 

matrix_corr <- round(cor(data_corr, method = "spearman"), 1)
```
```{r}
corrplot(matrix_corr, type = "lower", method = "ellipse", order = "alphabet")
```

## Results

We will next build a set of models to investigate our research question, documenting our decisions. 

### Model 1

Based on our exploration above, our dependent variable is *crime*, represented by the variable `crmrte`.

For our first model, we aimed to evaluate the relationship between crime policy and crime. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(acj) <- log(`prbarr` * `prbconv` * `prbpris`) = log(`prbarr`) + log(`prbconv`) + log(`prbpris`) 

(2) log(`avg_sen`)

However, `acj` and not `avg_sen` had a statistically significant relationship to crime rate, thus we removed `avg_sen` to limit our total number of variables in the model (to avoid overfitting). We already have important variables related to crime policy, thus including `avgsen` is superfluous.

### Data Transformations

We decided to take the log of `acj` and `crmrte` to improve our model fit and interpretability. With these variables logged, they can be interpreted as relative increases in crime rate and assertiveness of the criminal justice system (rather than their value alone). 

```{r}
ggplot(data = data2, aes(x = acj, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Assertiveness of Criminal Justice System") + 
  xlab("Crime Rate") + ylab("Assertiveness of Criminal Justice System") +
  geom_smooth(method='lm', formula= y~x)
```

```{r}
ggplot(data = data2, aes(x = log(acj), y = log(crmrte))) +
  geom_point() +
  theme_minimal()  +
  ggtitle("Relationship between Log Crime Rate and Log Assertiveness of Criminal Justice System") + 
  ylab("Log Crime Rate") + xlab("Log Assertiveness of Criminal Justice System") +
  geom_smooth(method='lm', formula= y~x)
``` 

Here, we compare the use of arrest, conviction, and sentencing rates (assertiveness measure) with just arrest and conviction rates. Model fit appears to be better without the use of sentencing rates. The sentencing rate variable had relatively lower variation, which would limit the amount of variation it could explain in crime rate (and thus, how much explanatory power it had). 

```{r}
ggplot(data = data2, aes(x = log(crmrte), y = log(prbarr) + log(prbconv))) +
  geom_point() +
  theme_minimal()  +
  ggtitle("Relationship between Log Crime Rate and Log Arrest & Log Conviction Rates") + 
  ylab("Log Crime Rate") + xlab("Log Arrest + Log Conviction Rates") +
  geom_smooth(method='lm', formula= y~x)
``` 

### Relationship between Crime Rate and Assertiveness of the Criminal Justice System

Approximately 30% of the variation in crime rate can be explained by the assertiveness of the criminal justice process. Increasing the ratio of criminal justice practices to crime (increasing arrests, convictions, and sentencing) could help decrease the crime rate. 

```{r, warning=F, error=F}
mod1 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv), data = data2)
stargazer(mod1, type = "text")
```

### Assumptions 

#### Zero Conditional Mean

We are close to meeting the zero conditional mean assumption, based on our plots below. Some residuals diverse at the extremes of our distribution, however for the most part they are similar. 

**Heteroskadiscity:**

We can see from the plots below that we do not veer far from homoskedasticity -- some points in our Normal Q-Q plot appear to diverge when they are at the extremes, but for the most part they are similar in their variance.

```{r}
plot(mod1)
```

#### Normality of Errors

```{r}
ggplot(data = data2, aes(x = mod1$fitted.values, y = log(data2$crmrte))) + 
  geom_point() +
  ggtitle("Model 1: Actual vs. Predicted Values") +
  theme_minimal() +
  xlab("Predicted Log Crime Rate") + ylab("Actual Log Crime Rate") +
  geom_smooth(method='lm', formula= y~x)
```



#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model (observation #24). 

```{r}
ols_plot_resid_lev(mod1)
```

The outlier with leverage was in county #11. 

```{r}
data2 %>% filter(county == 11)
```

We can see that without this observation, we now explain slightly more of the variation (~44% compared to ~40%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 11)

mod1_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv), data = data2_out)
stargazer(mod1, mod1_out, type = "text")
```

Although eliminating the outlier has improved our fit and how well our model meets our assumptions. However, we did not spot any data quality issues in this county during our exploratory data analysis thus we cannot claim that it has a data quality issue. 

```{r}
plot(mod1_out)
```

```{r}
ols_plot_resid_lev(mod1_out)
```

### Model 2

For our second model, we aimed to evaluate the relationship between crime policy, economic policy, and crime.

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(`prbarr`) + log(`prbconv`)

*Economic policy* 

(2) log(`wtrd`) 

(3) log(`wfed`) 

(4) log(`taxpc`)

All of the above variables had statistically significant relationships to crime rate, except `wtrd`. Additionally, this variable's relationship was not in the expected direction - rather than seeing higher wages related to lower levels of crime (as we predicted), we saw that higher wages were related to higher levels of crime. Since we cannot interpret this variable and our measured relationship of it to crime rate could be due to chance, we decided to omit this variable. This means that we are not able to make any recommendations regarding minimum wage from these data. 

### Data Transformations

We decided to take the log of our two wage variables and the tax revenue variable to show the relationship between relative changes in these to crime (rather than the absolute number). This improves our interpretability since we are making policy recommendations around changing these relative to the existing levels in these counties. 

We can see that we have a somewhat linear relationship between trade industry wages and crime. 

```{r}
ggplot(data = data2, aes(x = wtrd, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Trade Wages") + 
  xlab("Crime Rate") + ylab("Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging these wages.

```{r}
ggplot(data = data2, aes(x = log(wtrd), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Trade Wages") + 
  xlab("Crime Rate") + ylab("Log Trade Wages") +
  geom_smooth(method='lm', formula= y~x)
```

We can see that we have a relatively clear linear relationship between federal wages and crime. 

```{r}
ggplot(data = data2, aes(x = wfed, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Federal Employee Wages") + 
  xlab("Crime Rate") + ylab("Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging these wages - especially in terms of variability. One can see that the spread appears more even across different logged wages (compared to wages). 

```{r}
ggplot(data = data2, aes(x = log(wfed), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Federal Employee Wages") + 
  xlab("Crime Rate") + ylab("Log Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

We can see that we do not have a very linear relationship between tax per capita and crime. Many of the points cluster together, leaving little variability with which to predict crime. 

```{r}
ggplot(data = data2, aes(x = taxpc, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Tax Per Capita") + 
  xlab("Crime Rate") + ylab("Tax Per Capita") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging tax per capital. 

```{r}
ggplot(data = data2, aes(x = log(taxpc), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Tax Per Capita") + 
  xlab("Crime Rate") + ylab("Log Tax Per Capita") +
  geom_smooth(method='lm', formula= y~x)
```

### Relationship between Crime Rate and Assertiveness of the Criminal Justice System

Approximately 60% of the variation in crime rate can be explained by the assertiveness of the criminal justice process and with economic development factors. Increasing the ratio of criminal justice practices to crime (increasing arrests and convictions) along with efforts towards increasing the economic viability of a region (and thus increasing wages, cost of living, and tax revenue) could help decrease the crime rate. These two efforts should be made *together*, as there are additive effects of both. One can see that a model with only the criminal justice related variables (column 1) explains about 40% of the variance by itself. A model with only economic factors (column 2) explains about 33% of the variance. Together, they explain 59%. 

```{r, warning=F, error=F}
mod2a <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc), data = data2)

mod2b <- lm(log(crmrte) ~ log(wfed) +log(taxpc), data = data2)
stargazer(mod1, mod2b, mod2a, type = "text")
```

### Assumptions 

#### Zero Conditional Mean

We are close to meeting the zero conditional mean assumption, based on our plots below. Some residuals diverse at the extremes of our distribution, however for the most part they are similar. 

**Heteroskadiscity:**

We can see from the plots below that we do not veer far from homoskedasticity -- some points in our Normal Q-Q plot appear to diverge when they are at the extremes, but for the most part they are similar in their variance.

```{r}
plot(mod2a)
```

#### Normality of Errors

```{r}
ggplot(data = data2, aes(x = mod2a$fitted.values, y = log(data2$crmrte))) + 
  geom_point() +
  ggtitle("Model 2: Actual vs. Predicted Values") +
  theme_minimal() +
  xlab("Predicted Log Crime Rate") + ylab("Actual Log Crime Rate") +
  geom_smooth(method='lm', formula= y~x)
```



#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model (observation #24). 

```{r}
ols_plot_resid_lev(mod2a)
```

Once again, county #11 appears to be an outlier with leverage.  

```{r}
data2 %>% filter(county == 11)
```

We can see that without this observation, we now explain slightly more of the variation (~62% compared to ~59%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 11)

mod2_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc), data = data2_out)
stargazer(mod2a, mod2_out, type = "text")
```

Although eliminating the outlier has improved our fit, it does not appear to have greatly affected how closely we meet our assumptions. Additionally, although this outlier has leverage, we do not have any reason to suspect that it has data quality issues. 

```{r}
plot(mod2_out)
```

```{r}
ols_plot_resid_lev(mod2_out)
```


### Model 3

For our third model, we aimed to evaluate the robustness of the relationship between crime policy, economic policy, and crime when controlling for key contextual factors. We aimed for parsimony to avoid over-fitting, and following the one to ten rule we should aim to have about one independent variable for every 10 observations. We have 90 observations, meaning that ~9 independent variables is a good number to aim for. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1/2) log(`prbarr`) + log(`prbconv`)

*Economic policy* 

(3) log(`wfed`) 

(4) log(`taxpc`)

*Contextual variables* 

(5) `density` - higher density means more opportunity for crime to be committed and reported; denser areas also tend to have more economic activity and so this variable could be related to wages 

(6) `mix` - different types of crime in an area may affect the assertiveness of the criminal justice system; a community may see more arrests and convictions in response to more frightening face-to-face crimes in an area

(7) `pctymle` - young males are more likely to commit crimes and thus areas with higher portions of this type of person may see more crime regardless of their policies 

(8) `pctmin80` - minorities are likelier to be arrested and convicted (and from a lot of research we know that this is not always related to their likelihood to commit a crime)

(9) `west` - Western North Carolina is potentially very different from other regions in North Carolina, as we noted at the beginning of this report. A different culture could relate to different crime patterns and relationships between crime and other factors. For example, this variable is highly related to `pctmin80` (Western counties have lower levels of minorities) and this could affect crime rate through differences in diversity. 

These were all statistically significant when we modelled them together, except for `mix` and `west`. For the sake of parsimony, we dropped these two from our third model. 

### Data Transformations

Many of the contextual variables (`density`, `mix`, `pctymle`) showed a pattern where most counties clustered closely together. Logging them did not improve their fit. This similar pattern may be an artefact of North Carolina being a small state that may have limited variation across counties on some elements (like density, crime patterns in terms of their type, and percent young male). 

#### Density

Taking the log of density improves its linear fit to crime rate. 

```{r}
ggplot(data = data2, aes(x = density, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Density") + 
  ylab("Crime Rate") + xlab("Density") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(density), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Density") + 
  ylab("Crime Rate") + xlab("Log Density") +
  geom_smooth(method='lm', formula= y~x)
```

Taking the log of mix reduced the linear fit with crime rate. 

```{r}
ggplot(data = data2, aes(x = mix, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Mix") + 
  ylab("Crime Rate") + xlab("Mix") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(density), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Mix") + 
  ylab("Crime Rate") + xlab("Log Mix") +
  geom_smooth(method='lm', formula= y~x)
```

#### Percent Young Male

```{r}
ggplot(data = data2, aes(x = pctymle, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Percent Young Male") + 
  ylab("Crime Rate") + xlab("% Young Male") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(pctymle), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Percent Young Male") + 
  ylab("Crime Rate") + xlab("Log % Young Male") +
  geom_smooth(method='lm', formula= y~x)
```

#### Percent Minority

The percent minority in 1980 does not have a strong relationship to crime. Counties appear to have a wide range of this variable and there is a slight positive relationship to crime rate. Logging this variable did not improve the fit. 

```{r}
ggplot(data = data2, aes(x = pctmin80, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Percent Minority") + 
  ylab("Crime Rate") + xlab("% Minority") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(pctmin80), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Percent Minority") + 
  ylab("Crime Rate") + xlab("Log % Minority") +
  geom_smooth(method='lm', formula= y~x)
```

#### West

Western counties appear to have lower rates of crime than others. 

```{r}
ggplot(data = data2, aes(x = west, crmrte)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Western Location") + 
  ylab("Crime Rate") + xlab("West (0 = no; 1 = yes)") 

```

### Relationship between Crime Rate and Assertiveness of the Criminal Justice System

Including factors related to the demographics of an area (density; percent young males; and percent minorities) increases our ability to predict crime. Thus, these factors should be part of policy considerations we recommended above. Higher rates of any of these are related to higher levels of crime. Although we would not sensibly or ethically craft policies to affect these, policies designed to impact crime or economic policy should be mindful of the fact that some areas will naturally have higher crime and may be more challenging to impact. 


```{r, warning=F, error=F}
mod3 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2)
stargazer(mod3, type = "text")
```

### Assumptions 

#### Zero Conditional Mean

We are close to meeting the zero conditional mean assumption, based on our plots below. Some residuals diverse at the extremes of our distribution, however for the most part they are similar. 

**Heteroskadiscity:**

We can see from the plots below that we do not veer far from homoskedasticity -- some points in our Normal Q-Q plot appear to diverge when they are at the extremes, but for the most part they are similar in their variance.

```{r}
plot(mod3)
```

#### Normality of Errors

```{r}
ggplot(data = data2, aes(x = mod3$fitted.values, y = log(data2$crmrte))) + 
  geom_point() +
  ggtitle("Model 3: Actual vs. Predicted Values") +
  theme_minimal() +
  xlab("Predicted Log Crime Rate") + ylab("Actual Log Crime Rate") +
  geom_smooth(method='lm', formula= y~x)
```



#### Outliers

One can see that there are two potential outliers that also have leverage on our model (observations #6 & 25). We also have one borderline case.  

```{r}
ols_plot_resid_lev(mod3)
```

Once again, county #11 appears to be an outlier with leverage. County #55 is new in terms of outliers. However, we previously flagged County 185 as potentially having data quality issues.  

```{r}
data2 %>% filter(county %in% c(11, 55, 185))
```

We can see that without this observation, we now explain a lot more of the variation (~78% compared to ~72%). Additionally, `taxpc` is no longer significant.  

```{r, warning = F, error = F}
data2_out2 <- data2 %>% filter(county != 11) %>% filter(county != 55) %>% filter(county != 185)

mod3_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2_out2)
stargazer(mod3, mod3_out, type = "text")
```

Although eliminating the outlier has improved our fit and changed the strength of the relationship with `taxpc`, it does not appear to have greatly affected how closely we meet our assumptions. Additionally, although counties #11 & 55 have leverage, we do not have any reason to suspect that they have data quality issues. 

```{r}
plot(mod3_out)
```

What if we only eliminate the county with data quality issues? Now our model fit is slightly less improved (73% compared to the original 72%). The variable `taxpc` is again signficant. 

```{r, warning = F, error = F}
data2_out3 <- data2 %>% filter(county != 185)

mod3_out2 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2_out3)
stargazer(mod3, mod3_out, mod3_out2, type = "text")
```

For now, we will leave all of these in our model since #11 & 55 do not appear to have data quality issues and #185 does not change our model signficantly. 


### Model 4

For our fourth model, we aimed to evaluate the robustness of the relationship between crime policy, economic policy, and crime when controlling for many contextual factors. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1/2) log(`prbarr`) + log(`prbconv`)

*Economic policy* 

(3) log(`wfed`) 

(4) log(`taxpc`)

*Contextual variables* 

(5) `density` - higher density means more opportunity for crime to be committed and reported; denser areas also tend to have more economic activity and so this variable could be related to wages 

(6) `mix` - different types of crime in an area may affect the assertiveness of the criminal justice system; a community may see more arrests and convictions in response to more frightening face-to-face crimes in an area

(7) `pctymle` - young males are more likely to commit crimes and thus areas with higher portions of this type of person may see more crime regardless of their policies 

(8) `pctmin80` - minorities are likelier to be arrested and convicted (and from a lot of research we know that this is not always related to their likelihood to commit a crime)

(9) `west` - Western North Carolina is potentially very different from other regions in North Carolina, as we noted at the beginning of this report. A different culture could relate to different crime patterns and relationships between crime and other factors. For example, this variable is highly related to `pctmin80` (Western counties have lower levels of minorities) and this could affect crime rate through differences in diversity. 

(10) log(`wtrd`) 

(11) log(`prbpris`) 

(12) `central`

(13) `urban`

These were all statistically significant when we modelled them together, except for `mix` and `west`. For the sake of parsimony, we dropped these two from our third model. 

### Data Transformations


### Relationship between Crime Rate and Assertiveness of the Criminal Justice System

We can see that none of our additions improved our model fit. 

```{r, warning=F, error=F}
mod4 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80 + log(wtrd) +  log(`prbpris`) + avgsen + central + urban, data = data2)
stargazer(mod4, type = "text")
```

### Assumptions 

#### Zero Conditional Mean

We are close to meeting the zero conditional mean assumption, based on our plots below. Some residuals diverse at the extremes of our distribution, however for the most part they are similar. 

**Heteroskadiscity:**

We can see from the plots below that we do not veer far from homoskedasticity -- some points in our Normal Q-Q plot appear to diverge when they are at the extremes, but for the most part they are similar in their variance.

```{r}
plot(mod4)
```

#### Normality of Errors

```{r}
ggplot(data = data2, aes(x = mod4$fitted.values, y = log(data2$crmrte))) + 
  geom_point() +
  ggtitle("Model 4: Actual vs. Predicted Values") +
  theme_minimal() +
  xlab("Predicted Log Crime Rate") + ylab("Actual Log Crime Rate") +
  geom_smooth(method='lm', formula= y~x)
```


#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model. 

```{r}
ols_plot_resid_lev(mod4)
```

Once again, county #55 appears to be an outlier with leverage.  

```{r}
data2 %>% filter(county == 55)
```

We can see that without this observation, we now explain slightly more of the variation (~74% compared to ~72%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 55)

mod4_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80 + log(wtrd) +  log(`prbpris`) + avgsen + central + urban, data = data2_out)
stargazer(mod4, mod4_out, type = "text")
```

Although eliminating the outlier has improved our fit, it does not appear to have greatly affected how closely we meet our assumptions. Additionally, although this outlier has leverage, we do not have any reason to suspect that it has data quality issues. 

```{r}
plot(mod4_out)
```

```{r}
ols_plot_resid_lev(mod4_out)
```


### A Regression Table

Of our models, the third appears to be our best for predicting crime. We have included both crime and economic variables in the model, as well as some key contextual variables. 

```{r, warning=F, error=F}
stargazer(mod1, mod2a, mod3, mod4, type = "text")
```

# A Discussion of Omitted Variables
Identify what you think are the 5-10 most important omitted variables that bias results you care about. For each variable, you should estimate what direction the bias is in. If you can argue whether the bias is large or small, that is even better. State whether you have any variables available that may proxy (even imperfectly) for the omitted variable. Pay particular attention to whether each omitted variable bias is towards zero or away from zero. You will use this information to judge whether the effects you find are likely to be real, or whether they might be entirely an artifact of omitted variable bias.

**Education levels of the populace** 

Although we may be able to proxy education levels by looking at which region a county is in, we would not have a very good measure of overall education levels. A county can have a university, but still have low education levels overall. One would predict that education is negatively related to crime - the more educated, the less crime one would expect to see. Education levels would be positively correlated to tax per capita, wages, and density. It's unclear what direction the relationship would be to the distribution of young males or minorities, or to criminal justice practice. Thus, one would guess that the bias would be towards zero (positive relationship to tax per capital * positive relationship to wages * positive relationship to density * negative relationship to crime = negative).  

**Public works & investments** 

One would predict that, like education, the relationship between and public works and investments would be negative. More nonprofits and community organizations could strengthen ties in communities, thus lowering crime rates. This factor would probably be inversely related to arrest and conviction rates, since a more punitive society may invest less in rewards to the community. It would probably be positively related to wages and tax per capita, since more money means more funds available to invest back. It would also be positively related to density, since cities often have a lot more opportunities to create and access these investments (think of the museums in NYC). The direction of the relationship to the distribution of young males or minorities is unclear. Thus, one would guess that the bias would be towards zero (negative relationship to arrest rate * negative relationship to conviction rate * positive relationship to tax per capita * positive relationship to wages * positive relationship to density * negative relationship to crime = negative).  

**Individual Health (mental and physical)** 

Individual health - both mental and physical - would likely be related to less crime. The healthier you are, the more mental and physical resources you have to solve problems in ways that do not necessitate committing a crime. One would expect this factor to be positively related to wages and tax per capita but negatively related to percent young male (they take more risks) and minorities (racism has negative impacts on health). The direction of the relationship to arrest or conviction rates is unclear. The direction of the relationship to density could go either way - people in cities may have access to better and more accessible health care, but live in a more polluted environment. Thus, one would guess that the bias would be towards zero (positive relationship to tax per capital * positive relationship to wages * negative relationship to percent young male * negative relationship to percent minority * negative relationship to crime = negative).

**Community Health**

Community health is about the cohesiveness of a community. If you have good supports in your community, you would be less likely to commit crime. The relationship to wages and tax per capita is unclear - communities at many income levels can have loving, supportive relationships. It is unclear what the relationship would be to percent young males or minorities or density. This factor would probably be inversely related to arrest and conviction rates, since a more loving, cohesive community may be less likely to turn to punishment. Thus, one would guess that the bias would be towards zero (negative relationship to arrest rate * negative relationship to conviction rate * negative relationship to crime = negative).

**Inequality** 

We only had measures of average wage by sector in each county. We don't have information about the distribution within and across these sectors. With more inequality, one would expect to see more crime - more anger & therefore reason to commit crime combined with greater levels of deprivation & need to commit crime would naturally lead to more crime. The direction of the relationship between inequality and any of our independent variables in our final model is unclear - thus it is difficult to tell what direction this bias would be in. 

# A Conclusion

Our analyses suggest that having a two-pronged approach - deterrants (like arrests and convictions) combined with economic development - may be important policy efforts to take when it comes to preventing crime. None of these effects can be interpreted as causal as we conducted a purely correlational analysis. We did not have variation across time for conducting a difference-in-differences test or large changes in policy that would help with regression discontinuity. However, our findings can be used to hone in on future policy changes that can be evaluated with causal analyses. 

These efforts do not detract from each other, but are close to equal in their importance. Other factors outside of policy - such as density and the distribution of young males or minorities - are also important when it comes to crime. Communities with higher rates of either may warrant additional resources to prevent crime. 

Some policies could include: 

* Improved training and hiring of criminal justice staff, that could improve the ratio of arrest and convictions to crimes. Making sure that those who commit crimes are caught seems to be important in deterring crime - whether by keeping people off the streets, rehabilitating them, or scaring others away from commiting crimes 

* Targeting economic development policies that can bring higher wages and improve the quality of life in a region. We have seen that higher tax per capita and higher wages of federal employees (which are determined in part by the cost of living) are important when it comes to crime. One could start a campaign to bring a well-funded federal effort to the region that would bring higher paying federal employees. Or, you could try to attract higher-paying corporations that would then increase wages for all. You could also increase taxes on some to bring in more revenue streams. 

Both of these streams (criminal and economic reform) could be folded under the umbrella of investing more in one's community. It is not enough to punish and deter - one needs to also get at the root causes of crime and why someone may be compelled to commit a crime in the first place. 




