---
title: 'Lab 3: Draft Report'
author: "Priscilla Burity, Oscar Linares, Alissa Stover"
date: "3/31/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Which policies are most promising in reducing the crime rate: economic (e.g., increasing wages) or classic criminal deterrants (e.g., increasing arrests) (or both together)? What is the influence of contextual factors (e.g., characteristics of a population) on these relationships? 

We have a cross-section data on crime statistics and indicators that can influence crime rates for a selection of counties in North Carolina. Our task is to examine the data to help a political campaign understand the determinants of crime and generate policy suggestions that are applicable to local government.

We will be modeling our target variable (crime rate, `crmrte`) using candidates of policy variables that represent different economic and criminal justice indicators that can be influenced via various policy decisions taken by our candidate. 

We also include a range of contextual variables that help to control for local characteristics (for example, demographic and geographic indicators) that could influence crime and the efficacy of these policies but are not actionable by a local government. We aim to produce a model robust to these, that can help guide policy decisions and help reduce crime across these counties. 

## Actionable Variables

### Economic Policy

These variables describe things we could affect with economic policy. For example, if increased wages are found to be related significantly to crime, our candidate could consider strategies such as raising the minimum wage in an attempt to lower crime. These variables speak to wages for different sectors.

- `wcon`:	weekly wage, construction
- `wtuc`:	wkly wge, trns, util, commun
- `wtrd`:	wkly wge, whlesle, retail trade
- `wfir`:	wkly wge, fin, ins, real est
- `wser`:	wkly wge, service industry
- `wmfg`:	wkly wge, manufacturing
- `wsta`:	wkly wge, state employees
- `wloc`:	wkly wge, local gov emps

The following federal wage variable will be grouped with other wages in much of our analyses for easier comparison in terms of data quality, however it is not related to minimum wage policy but rather to cost of living. Federal employee wages are adjusted by cost of living in a locality (https://www.opm.gov/policy-data-oversight/pay-leave/salaries-wages/fact-sheets/); thus, this variable could also be used as a marker of cost of living in a county. Cost of living is related to overall wealth in an area and its overall econimic development. If we find a relationship between federal wages and crime, one could predict that changing the economic development of a region may affect crime. However, this variable is confounded by the types of federal workers in an area, since federal wages are determined by a grade system that is tied to the nature of the position and time in position. A specific effort that could affect both of these - bringing more economic development and the right types of high-earning federal employees - would be to attract federal jobs to the area that have higher wages. 

- `wfed`:	wkly wge, fed employees

The following variable represents tax policy, but it could also reflect overall wealth in a county. Holding tax rate constant, counties with more wealth will have more tax revenue and vice versa. Holding wealth constant, changing the tax rate would change the tax revenue per capita. A better measure of tax policy would be the actual tax rates in each county. Since the impact of policy on this variable is confounded by the overal wealth in a county, the causal relationship would be difficult to interpret. Plus, many taxes are set at the state and federal level; the level of variation at the county-level may be limited for a small state like North Carolina. However, similar to `wfed`, this variable would be related to overall economic development. If we found a relationship between tax revenue and crime rate, it would make sense to try to change development efforts in order to affect crime.  

- `taxpc`:	tax revenue per capita 

### Criminal Justice Policy

These variables describe things we could affect with policy around the criminal justice system. Many of these policies are used as deterrants of crime, with the assumption being that people inherently understand that these punishments exist and as such will be less likely to commit crimes (in order to avoid the punishment). 

Please note that the 3 variables below are dubbed "probabilities". However, this is a misnomer as these variables are truly ratios (e.g., the ratio of arrests to crimes reported). We interpret these as measures of the assertiveness of the criminal justice system in a county. Ideally, we would want to strike a balance between arresting enough people to prevent criminal activity, and not arresting so many as to abuse the populace.

*Probability of arrest* 

If we find that the efficiency of arrest is associated with crime rate (with higher punishment associated with lower crime), we may consider stricter policies around police practice that encourage more arrests. 

- `prbarr`: number of convictions for every crime reported

*Probability of conviction* 

If we find that the efficiency of conviction is associated with crime rate (with higher conviction rates associted with lower crime), we may consider stricter policies around court practices so that more arrests lead to convictions.

- `prbconv`: number of convictions for every arrest made

*Probability of sentencing* 

If we find that the efficiency of sentencing is associated with crime rate (with higher sentencing rates associted with lower crime), we may consider stricter policies around judicial practices so that more convictions lead to sentencing. However, whether or not a given conviction carries a prison sentence depends on the severity of the crime. Thus, this variable may be more limited in its explanatory power than arrests and convictions. 

- `prbpris`: number of prison sentences for every conviction

*Severity of punishment* 

If we find that the severity of punishment is associated with crime rate (with greater severity associted with lower crime), we may consider stricter policies around judicial practices so that people receive more severe punishments if sentenced with a crime.

- `avgsen`:	avg. sentence, days 

*Number of policy officers per capita*

The number of police police officers per capita (measured by the variable `polpc`) is a very intuitive example of crime deterrant, as it in theory increases the likelyhood of being caught in the act of the crime and is also a measure of the capacity of the State to enforce the law.

As our job is to advise a policy maker, it’s very important that we feel safe about the causal interpretation of our findings. After all, the idea is to propose policies with a significant impact on people’s lives. Thus, for statistical rigor, we need to make sure that causal effects go one way (that we have reason to believe that our independent variables have causal effects on our dependent variables, and not the reverse). 

A cross section of counties though does not allow us to link police presence to crime rates with a causal interpretation. This is because one can expect the counties with larger crime rates in a year *t* would consider necessary to have a larger number of police officers on the streets in year *t*, which can imply a positive correlation between presence of police and crime rates that couldn't be read as police presence leading to *higher* crime rates. In fact, an increase in crime could result in an increase in the number of police officers *and* an increase in police officers could result in a decrease in crime. 

Thus, in this exercise we opted not to include police per capita as a candidate policy variable or covariate. 

## Contextual Variables

Since we are aiming to produce policy recommendations, the following variables may be helpful in our analysis but are not necessarily actionable. They describe county characteristics that provide important information about how policies could work across different county contexts.

### Types of Crime 

- `mix`:	offense mix: face-to-face/other

### Demographics 

The following variables describe different aspects of the people that reside in these counties.

*Urban/rural dwellers* 

- `density`:	people per sq. mile 

*Minority status*

- `pctmin80`:	perc. minority, 1980 

*Gender & Age* (young males are more likely to enter the criminal justice system)

- `pctymle`:	percent young male 


### Geography

The following variables identify which region of the state each county is in, with the assumption that counties cluster geographically in terms of their culture and other characteristics we are not explicitly measuring. Western North Carolina is along the Appalachian mountains and seems qualitatively different from much of the rest of the state. For example, it has much fewer universities compared to the rest of the state (you can see this through a quick Google Maps search). Thus, we will lean towards using `west` over the other geographical variables (especially since `urban` could be highly related to `density`). 

- `west`:	=1 if in western N.C.
- `central`:	=1 if in central N.C.
- `urban`:	=1 if in SMSA

# A Model Building Process

## Data Cleaning

### Environment

#### Import Libraries 

```{r, message=FALSE}
library(tidyverse) # for data import, manipulation, viz
library(corrplot) # for correlation matix
library(stargazer) # visualize model fit
library(skimr) # generate summary statistics
library(car) # statistics 
library(lmtest) # linear modeling
library(olsrr) # evaluating OLS regression 
library(sandwich) # correcting heteroskedasticity violation
```

#### Import Data

We are importing a CSV file into R. The argument `n_max` is set to the number of rows of data we are provided (91). 

```{r, message=FALSE}
data <- read_csv("crime_v2.csv", n_max = 91)
```

#### Exploratory Data Analysis: File-level data checks

Let's take a visual check of our data. We have 91 observations and 25 columns. This is what we expected based on our original csv file. 

```{r}
head(data)
dim(data)
```

All 25 columns are numeric. 

```{r}
str(data)
```

##### The Where & When of these data 

It looks like we usually have 1 observation for each of the counties -- except for county #193. We do not have any missing values in the `county` column. 

```{r}
length(unique(data$county))
table(data$county, useNA = "always")
```

Having 2 observations for county #193 is surprising. We can see that these rows appear to be exact duplicates. 

```{r}
data %>% filter(county == 193)
```

We can see that `year` appears to be constant (which is expected). We also don't have any missing values for this column. We can drop this column in future data processing steps. 

```{r}
table(data$year, useNA = "always")
```

Based on the explorations above, we create a new dataframe that has dropped our superfluous column (`year`) and the superfluous row (duplicate of county 193). 

```{r}
data2 <- data %>% distinct() %>% select(-year)
```

Now we have 90 rows and 24 variables, with one observation per county.

```{r}
length(unique(data2$county)) == length(data2$county) # check 1 obs per county
str(data2) # examine dataframe 
```

##### Missingness & Data Validity

We can see from the following table that we don't have any missing data, which is good! 

However, we can see by the max for each column (p100), that we appear to have a few values that are extremely high. 

For example, one county has an average sentence length (`avgsen`) of 20.7 years.

```{r}
data2_skim <- skim(data2)
data2_skim
```

##### Preserve Clean Data

Save our clean dataset

```{r}
save(data2, file = "data_clean.rda")
```

## Exploratory Data Analysis: Variable-level data checks

### Dependent Variable: Crime Rate

Barring extreme issues with `crmrte` (crime rate), we would recommend focusing on this as our target variable, as what matters most to people is the level of crime around them every day. The justice system is used to enact justice and to deter crimes. Justice is important, and making sure that those that commit crimes are punished is one way of going about achieving a just and safer society. Of course, justice is a moral issue and can be seen from many different lenses, but the data we have in hands do not allow us to go in depth in this issue. 

We don't have enough data points to make a strong claim about the distribution of `crmtre`, however it appears to be approximately normal (with a slight skew which could be an artifact of not having any values below 0). 

```{r}
ggplot(data2, aes(crmrte)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  xlab("Crime Rate") +
  ylab("Count") +
  ggtitle("Histogram of Candidate Dependent Variable: Crime Rate")
```

From our boxplot we can see that most values are between 0.02 and 0.04, however we do have a range of values including some that are close to 0.10. We don't appear to have any spurious values. 

```{r}
ggplot(data2, aes(y = crmrte)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Crime Rate") +
  ggtitle("Boxplot of Candidate Dependent Variable: Crime Rate")
```

Our preferred variable (`crmrte`) appears to be of sufficient quality for our analyses and thus we will use it as our dependent variable. 

### Independent Variables 

#### Actionable Variables

##### Economic Policy

One of the things one can affect with economic policy are wages. In North Carolina in 1987, the minimum hourly wage was \$3.35. Given that most people work 40 hours per week, the minimum hourly wage would be \$134 per week.

In the plots below, we plot weekly wage by sector with a red vertical line indicating this minimum weekly wage.

We can see that `wrtd` (Wholesale/Retail/Trade) wages are closest to the minimum wage - which would suggest that wages in this sector would be most influenced by changes to the minimum wage. The next sector that would be influenced is `wser` (Service Industry). However, `wser` appears to have an outlier (someone is earning over $2000 weekly, which is well above the weekly wages for any other sector as well as `wser` in other counties). Additionally, `wser` may be a messier measure of income because many in that industry work in restaurants and make additional income through tips - however, some people may not report this income. Meanwhile, those working in Wholesale/Retail/Trade may be more likely to work in places like malls, which would not provide additional income as tips and would be a better measure of low-income wages. 

```{r}
# format data for subplots 
data_ep_long <- data2 %>% 
  select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc) %>%
  gather(key = var, value = value)

data_ep_long$var <- factor(data_ep_long$var,
                           levels = c("wcon", "wtuc", "wtrd", "wfir",
                                      "wser", "wmfg", "wfed", "wsta", "wloc"),
                           labels = c("Construction", 
                                      "Utilities/Transport", 
                                      "Wholesale/Retail/Trade",
                                      "Finance/Insurance/Real Estate",
                                      "Service Industry",
                                      "Manufacturing",
                                      "Federal Emp",
                                      "State Emp",
                                      "Local/State Emp"))
```

```{r}
# plot hist subplots
ggplot(data_ep_long, aes(value)) +
  geom_histogram(bins = 50) +
  facet_wrap(~var) +
  theme_minimal() +
  ggtitle("Histogram of Economic Policy Variables: Wages by Sector") +
  geom_vline(aes(xintercept = 134, color = "red")) +
  theme(legend.position = "none")

```

```{r}
# plot boxplot subplots 
ggplot(data_ep_long, aes(x = var, y = value)) +
  geom_boxplot() +
  coord_flip() + 
  theme_minimal() +
  xlab("") + 
  ylab("Weekly Wage") +
  ggtitle("Boxplot of Economic Policy Variables: Wages by Sector")+
  geom_hline(aes(yintercept = 134, color = "red")) +
  theme(legend.position = "none")

```


One can see from the plot below that `wser` for county #185 appears to be an incorrect value - all of the other sector wages are close to the mean for all counties, except for this one (which is about 10x the mean). 

```{r}
data_wages <- data2 %>% select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc)%>% colMeans() %>% as_tibble() %>% mutate(names = c("wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc"), Group = "Mean")

wages_185 <- data2 %>%filter(wser == max(wser)) %>% select(wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc) %>% colMeans() %>% as_tibble() %>% mutate(names = c("wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc"), Group = "County #185")

wages <- bind_rows(data_wages, wages_185)

ggplot(data = wages, aes(x = names, y = value, color = Group)) + 
  geom_point() +
  xlab('Wage Category') +
  ylab('Wage') +
  theme_minimal() +
  ggtitle("Comparing Wages by Sector between County 185 and Other Counties")

```

##### Criminal Justice Policy

As noted above, we combined our 3 core criminal justice policy variables (`prbarr`, `prbconv`, `prbpris`) into one measure of the assertiveness of the criminal justice system. It is a combined measure of the ratio of arrests, convictions, and prison sentences to crimes. It could be interesting to the policy maker to check how these criminal justice indicators work together preventing crime.    

First, we evaluated the quality of these 3 variables separately. 

```{r}
# format data for subplots 
data_cj_long <- data2 %>% select(prbarr, prbconv, prbpris) %>%
  gather(key = var, value = value)

# convert to factor for subplot labelling 
data_cj_long$var <- factor(data_cj_long$var, labels = c("Prob Arrest", "Prob Convicted", "Prob Sentenced"))
```

We can see that we have different distributions for these, but that they all could be considered somewhat close to normal, with different levels of skew. 

```{r}
# plot hist subplots
ggplot(data_cj_long, aes(value)) +
  geom_histogram(bins = 50) +
  facet_grid(~var) +
  theme_minimal() +
  ggtitle("Histogram of Criminal Justice Policy Variables: Assertiveness Variables")
```

From our boxplots we can see that we have some extreme values for efficiency of conviction, with multiple counties having a rate above 1. As these are not true probabilities, but rather ratios, these values are not necessarily spurious. However, they imply that for each arrest, there could be multiple convictions - one county has over 2 convictions for every arrest. Only 1 county has more than 1 arrest per crime committed - most counties in fact have a much lower rate (and only put someone under arrest for about one-quarter of crimes). When it comes to setencing, counties show a much lower spread and cluster around 1 conviction resulting in a prison sentence for every 1 conviction that does not. 

```{r}
# plot boxplot subplots 
ggplot(data_cj_long, aes(x = var, y = value)) +
  geom_boxplot() +
  theme_minimal() +
  xlab("") +
  ylab("Ratio") + 
  ggtitle("Boxplot of Criminal Justice Policy Variables: Assertiveness Variables")
```

Which county had the highest conviction rate? This is a county in central North Carolina. It has the highest percentage of minorities. It also appears to have the maximum wages in the service industry ($2177.068 compared to a median of \$253.12). 

```{r}
data2 %>% filter(prbconv == max(prbconv))
```

The fact that the conviction rate is so high seems odd for county #185. The other measures of assertiveness (for arrest and sentencing) seem close to the mean. That of conviction rate is astronimically high. 

```{r}
data_prb <- data2 %>% select(prbarr, prbconv, prbpris) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris"), Group = "Mean")

prb_185 <- data2 %>% filter(prbconv == max(prbconv)) %>% select(prbarr, prbconv, prbpris) %>% colMeans() %>% as_tibble() %>% mutate(names = c("prbarr", "prbconv", "prbpris"), Group = "County #185")

prbs <- bind_rows(data_prb, prb_185)

ggplot(data = prbs, aes(x = names, y = value, color = Group)) + 
  geom_point() +
  xlab('Criminal Justice Efficiency Category') +
  ylab('Efficiency') +
  theme_minimal() +
  ggtitle("Comparing Criminal Justice Efficiencies between County 185 and Other Counties")

```

The analysis above suggests that the data for this county #185 may have been corrupted - since it has values for `wser` and `prbconv` at levels much higher than the mean, whithout any explanation for why these might have both been elevated. We cannot know for sure as we do not have any other data with which to triangulate these records, thus in our modeling we will compare models with and without this county; if it does not affect the models, we can leave in this outlier. If it does significantly, we may remove it since the data do not seem correct. 

Our dataset has a second kind of criminal justice system indicator, namely the average sentence of convictions (in days). While the assertiveness indicators relate to the certainty of punishment (the extent to which criminals expect to get caught and face punishment), average sentence relates to the severity of the punishment. We can see that average sentence length follows an approximately normal distribution. We have one value well above the rest (above 20; most cluster below 15). 

```{r}
# plot hist
ggplot(data2, aes(avgsen)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  ggtitle("Histogram of Criminal Justice Policy Variables: Severity") +
  xlab("Average Sentence Severity")
```

```{r}
# plot hist
ggplot(data2, aes(y = avgsen)) +
  geom_boxplot() +
  theme_minimal() +
  ggtitle("Boxplot of Criminal Justice Policy Variables: Severity") +
  ylab("Average Sentence Severity")
```

The county with the highest average sentence is in Western North Carolina. It has the lowest percentage of minorities of any county (1.28%). Its density is in the lowest quartile of counties but its mix of face-to-face versus other crime is at the median. Although this county seems to be at an extreme, there is no reason to eliminate it from the data at this juncture. 

#### Contextual Variables

##### Types of Crime 

```{r}
# plot barplots
ggplot(data2, aes(mix)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  ggtitle("Histogram of Contextual Variables: Type of Crime") +
  xlab("offense mix: face-to-face/other")
```

##### Demographics 

*Urban/rural dwellers* 

```{r}
ggplot(data2, aes(density)) +
  geom_histogram(bins = 35) +
  theme_minimal() +
  xlab("Density") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Population Density")
```


```{r}
ggplot(data2, aes(y = density)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Density") +
  ggtitle("Boxplot of Contextual Variables: Population Density")
```

*Minority status*

```{r}
ggplot(data2, aes(pctmin80)) +
  geom_histogram(bins = 40) +
  theme_minimal() +
  xlab("% Minority") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Percentage Minority (1980)")
```


```{r}
ggplot(data2, aes(y = pctmin80)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("% Minority") +
  ggtitle("Boxplot of Contextual Variables: Percentage Minority (1980)")
```

- `pctmin80`:	perc. minority, 1980 

*Gender & Age* (young males are more likely to enter the criminal justice system)

```{r}
ggplot(data2, aes(pctymle)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  xlab("% Young Males") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Percentage Young Males")
```

```{r}
ggplot(data2, aes(y = pctymle)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("% Young Males") +
  ggtitle("Boxplot of Contextual Variables: Percentage Young Males")
```

- `pctymle`:	percent young male 

*Wealth*

```{r}
ggplot(data2, aes(taxpc)) +
  geom_histogram(bins = 25) +
  theme_minimal() +
  xlab("Tax Revenue Per Capita") +
  ylab("Count") +
  ggtitle("Histogram of Contextual Variables: Tax Revenue Per Capita")
```

```{r}
ggplot(data2, aes(y = taxpc)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Tax Revenue Per Capita") +
  ggtitle("Boxplot of Contextual Variables: Tax Revenue Per Capita")
```

- `taxpc`:	tax revenue per capita 

##### Geography

```{r}
# format data for subplots 
data_geo_long <- data2 %>% select(west, central, urban) %>%
  gather(key = var, value = value)
```

```{r}
# plot barplots
ggplot(data_geo_long, aes(value)) +
  geom_bar() +
  facet_grid(~var) +
  theme_minimal() +
  ggtitle("Barplot of Contextual Variables: Geography")
```

- `west`:	=1 if in western N.C.
- `central`:	=1 if in central N.C.
- `urban`:	=1 if in SMSA


### Relationships between Variables

One would expect variables within each of the above groups to be correlated to each other. 

This is a correlation matrix of all of our variables, using Spearman's due to the fact that not all variables follow a normal distribution. 

Our original policy variables for economic policy -- those related to wages -- appear to all be positively correlated with crime, which is opposite our original conjecture. Rather than seeing that increased wages are related to reduced crime, we are seeing that higher wages is related to higher crime. This unexpected direction could be explained in part by the strong positive relationship between the wage variables and `density` and between `crmrte` and `density`. Thus, we will not include the wage variables in early models - and it will be important to include additional variables to control for other influences that could be affecting the direction of this relationship. 

```{r}
data_corr <- data2 %>% select(-county) 

matrix_corr <- round(cor(data_corr, method = "spearman"), 1)
```
```{r}
corrplot(matrix_corr, type = "lower", method = "ellipse", order = "alphabet")
```

## Results

We will next build a set of models to investigate our research question, documenting our decisions. 

### Model 1

Based on our exploration above, our dependent variable is *crime*, represented by the variable `crmrte`.

For our first model, we aimed to evaluate the relationship between crime policy and crime. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(`prbarr`)

(2) log(`prbconv`) 

(3) log(`prbpris`) 

(4) log(`avg_sen`)


#### Data Transformations

We decided to take the log of the crime policy measures and `crmrte` to improve our model fit and interpretability. With these variables logged, they can be interpreted as relative increases in crime rate and assertiveness of the criminal justice system (rather than their value alone). 

#### Discussion on results

Coefficients on log(`prbpris`)  and log(`avg_sen`) were not statistically significant, thus we removed these variables to limit our total number of variables in the model (to avoid overfitting). We already have important variables related to crime policy, thus including them isn't crucial.

To make sure, here we also compare the use of arrest, conviction, and sentencing rates (assertiveness measures) with just arrest and conviction rates. Model fit appears to be better without the use of sentencing rates. The sentencing rate variable has relatively lower variation, which would limit the amount of variation it could explain in crime rate (and thus, how much explanatory power it has). 

```{r}
ggplot(data = data2, aes(x = log(crmrte), y = log(prbarr) + log(prbconv))) +
  geom_point() +
  theme_minimal()  +
  ggtitle("Relationship between Log Crime Rate and Log Arrest & Log Conviction Rates") + 
  ylab("Log Crime Rate") + xlab("Log Arrest + Log Conviction Rates") +
  geom_smooth(method='lm', formula= y~x)
``` 


Approximately 40% of the variation in crime rate can be explained by the assertiveness of the criminal justice process. Increasing the ratio of criminal justice practices to crime (increasing arrests, convictions, and sentencing) could help decrease the crime rate. 

```{r, warning=F, error=F}
mod1_1 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(prbpris) + log(avgsen), data = data2)
mod1_2 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(prbpris) , data = data2)
mod1 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) , data = data2)
stargazer(mod1_1, mod1_2, mod1, type = "text")
```


#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model (observation #24). 

```{r}
ols_plot_resid_lev(mod1)
```

The outlier with leverage was in county #11. 

```{r}
data2 %>% filter(county == 11)
```

We can see that without this observation, we now explain slightly more of the variation (~44% compared to ~40%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 11)

mod1_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv), data = data2_out)
stargazer(mod1, mod1_out, type = "text")
```

Eliminating the outlier has improved our fit. However, we did not spot any data quality issues in this county during our exploratory data analysis thus we cannot claim that it has a data quality issue. 

```{r}
ols_plot_resid_lev(mod1_out)
```

### Model 2

For our second model, we aimed to evaluate the relationship between crime and economic policies with crime.

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(`prbarr`) 
(2) log(`prbconv`)

*Economic policy* 

(3) log(`wtrd`) 

(4) log(`wfed`) 

(5) log(`taxpc`)


#### Data Transformations

We decided to take the log of our two wage variables and the tax revenue variable to show the relationship between relative changes in these to crime (rather than the absolute number). This improves our interpretability since we are making policy recommendations around changing these relative to the existing levels in these counties. 

We can see that we have a somewhat linear relationship between trade industry wages and crime. 

```{r}
ggplot(data = data2, aes(x = wtrd, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Trade Wages") + 
  xlab("Crime Rate") + ylab("Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging these wages.

```{r}
ggplot(data = data2, aes(x = log(wtrd), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Trade Wages") + 
  xlab("Crime Rate") + ylab("Log Trade Wages") +
  geom_smooth(method='lm', formula= y~x)
```

We can see that we have a relatively clear linear relationship between federal wages and crime. 

```{r}
ggplot(data = data2, aes(x = wfed, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Federal Employee Wages") + 
  xlab("Crime Rate") + ylab("Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging these wages - especially in terms of variability. One can see that the spread appears more even across different logged wages (compared to wages). 

```{r}
ggplot(data = data2, aes(x = log(wfed), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Federal Employee Wages") + 
  xlab("Crime Rate") + ylab("Log Federal Employee Wages") +
  geom_smooth(method='lm', formula= y~x)
```

We can see that we do not have a very linear relationship between tax per capita and crime. Many of the points cluster together, leaving little variability with which to predict crime. 

```{r}
ggplot(data = data2, aes(x = taxpc, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Tax Per Capita") + 
  xlab("Crime Rate") + ylab("Tax Per Capita") +
  geom_smooth(method='lm', formula= y~x)
```

The fit appears slightly improved after logging tax per capital. 

```{r}
ggplot(data = data2, aes(x = log(taxpc), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Tax Per Capita") + 
  xlab("Crime Rate") + ylab("Log Tax Per Capita") +
  geom_smooth(method='lm', formula= y~x)
```

#### Discussion on results

All of the above variables had statistically significant relationships to crime rate, except `wtrd`. Additionally, this variable's relationship was not in the expected direction - rather than seeing higher wages related to lower levels of crime (as we had conjectured), we saw that higher wages were related to higher levels of crime. 

Additionally, the estimated relationship between `taxpc` and `wfed` with `crmrte` was statistically significant but positive. This indicates crime rate is larger in wealthier counties.  

Since wage and wealth indicators and `density` for example are positively correlated, we thought that perhaps omitted variables could be influencing the direction of the estimated relationship between wage and wealth and `crmrte`.


```{r, warning=F, error=F}
mod2a <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wtrd) + log(wfed) +log(taxpc), data = data2)

mod2b <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wtrd) + log(wfed) +log(taxpc), data = data2)

stargazer(mod1, mod2b, mod2a, type = "text")
```

#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model (observation #24). 

```{r}
ols_plot_resid_lev(mod2a)
```

Once again, county #11 appears to be an outlier with leverage.  

```{r}
data2 %>% filter(county == 11)
```

We can see that without this observation, we now explain slightly more of the variation (~62% compared to ~59%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 11)

mod2_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc), data = data2_out)
stargazer(mod2a, mod2_out, type = "text")
```

Although eliminating the outlier has improved our fit and this outlier has leverage, we do not have any reason to suspect that it has data quality issues. 

```{r}
ols_plot_resid_lev(mod2_out)
```

In checking our OLS assumptions for model 2, we noted that we violated the homoskedasticity assumption. The variability of our residuals decrease as the value of our target variable (crime) increases. 

```{r, warning=F, error=F}
plot(mod2a, which = 3)
```

We can see that log(`taxpc`), `density`, and `pctymle` are no longer significant. Thus, we will only make conclusions regarding our crime policy variables, wages of federal employees, and percent minorities. However, we will continue to use these as we continue our modeling to examine how the inclusion of additional covariates  

```{r}
coeftest(mod2a, vcov. = vcovHC)
```

### Model 3

For our third model, we added density and other control variables to check if, controling for them, the apparent irrelevance of `wtrd` and unexpected positive coefficient estimate for log(`wfed`) would persist.

We also we aimed to evaluate the robustness of the relationship between crime policy, economic policy, and crime when controlling for key contextual factors. We aimed for parsimony to avoid over-fitting, and following the one to ten rule we should aim to have about one independent variable for every 10 observations. We have 90 observations, meaning that ~10 independent variables is a good number to aim for. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(`prbarr`)

(2) log(`prbconv`)

*Economic policy* 

(3) log(`wtrd`) 

(4) log(`wfed`) 

(5) log(`taxpc`)

*Contextual variables* 

(6) `density` - higher density means more opportunity for crime to be committed and reported; denser areas also tend to have more economic activity and so this variable could be related to wages 

(7) `mix` - different types of crime in an area may affect the assertiveness of the criminal justice system; a community may see more arrests and convictions in response to more frightening face-to-face crimes in an area

(8) `pctymle` - young males are more likely to commit crimes and thus areas with higher portions of this type of person may see more crime regardless of their policies 

(9) `pctmin80` - minorities are likelier to be arrested and convicted (and from a lot of research we know that this is not always related to their likelihood to commit a crime)

(10) `west` - Western North Carolina is potentially very different from other regions in North Carolina, as we noted at the beginning of this report. A different culture could relate to different crime patterns and relationships between crime and other factors. For example, this variable is highly related to `pctmin80` (Western counties have lower levels of minorities) and this could affect crime rate through differences in diversity. 


#### Data Transformations

Many of the contextual variables (`density`, `mix`, `pctymle`) showed a pattern where most counties clustered closely together. Logging them did not improve their fit. This similar pattern may be an artefact of North Carolina being a small state that may have limited variation across counties on some elements (like density, crime patterns in terms of their type, and percent young male). 

##### Density

Taking the log of density improves its linear fit to crime rate. 

```{r}
ggplot(data = data2, aes(x = density, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Density") + 
  ylab("Crime Rate") + xlab("Density") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(density), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Density") + 
  ylab("Crime Rate") + xlab("Log Density") +
  geom_smooth(method='lm', formula= y~x)
```

Taking the log of mix reduced the linear fit with crime rate. 

```{r}
ggplot(data = data2, aes(x = mix, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Mix") + 
  ylab("Crime Rate") + xlab("Mix") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(density), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Mix") + 
  ylab("Crime Rate") + xlab("Log Mix") +
  geom_smooth(method='lm', formula= y~x)
```

##### Percent Young Male

```{r}
ggplot(data = data2, aes(x = pctymle, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Percent Young Male") + 
  ylab("Crime Rate") + xlab("% Young Male") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(pctymle), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Percent Young Male") + 
  ylab("Crime Rate") + xlab("Log % Young Male") +
  geom_smooth(method='lm', formula= y~x)
```

##### Percent Minority

The percent minority in 1980 does not have a strong relationship to crime. Counties appear to have a wide range of this variable and there is a slight positive relationship to crime rate. Logging this variable did not improve the fit. 

```{r}
ggplot(data = data2, aes(x = pctmin80, crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Percent Minority") + 
  ylab("Crime Rate") + xlab("% Minority") +
  geom_smooth(method='lm', formula= y~x)

ggplot(data = data2, aes(x = log(pctmin80), crmrte)) +
  geom_point() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Log Percent Minority") + 
  ylab("Crime Rate") + xlab("Log % Minority") +
  geom_smooth(method='lm', formula= y~x)
```

##### West

Western counties appear to have lower rates of crime than others. 

```{r}
ggplot(data = data2, aes(x = west, crmrte)) +
  geom_col() +
  theme_minimal() +
  ggtitle("Relationship between Crime Rate and Western Location") + 
  ylab("Crime Rate") + xlab("West (0 = no; 1 = yes)") 

```

#### Discussion of results 

As the coefficient estimate on log(`wfed`) and log(`taxpc`) remained positive and the coefficient on `wtrd` remained statistically not significant even controlling for potential ommited variables, we believe we are not able to make any recommendations regarding economic policies in general based on the avaiable dataset. 

Still, including factors related to the demographics of an area (density; percent young males; and percent minorities) increases our ability to predict crime. Thus, these factors should be part of policy considerations we recommended above. Higher rates of any of these are related to higher levels of crime. Although we would not sensibly or ethically craft policies to affect these, policies designed to impact crime or economic policy should be mindful of the fact that some areas will naturally have higher crime and may be more challenging to impact. 

These were all statistically significant when we modelled them together, except for `mix` and `west`. For the sake of parsimony, we dropped these two from our third model. As the avaiable datase does not seen to support the significance of a minimum wage policy, also dropped log(`wtrd`). We decided to keep log(`wfed`) and log(`taxpc`) and read them as controls for county wealth, rather than policy variables per se.    


```{r, warning=F, error=F}
mod3a<- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wtrd) + log(wfed) +log(taxpc) + density + pctymle + pctmin80 + mix + west, data = data2)
mod3<- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2)
stargazer(mod3a, mod3, type = "text")
```

#### Outliers

One can see that there are two potential outliers that also have leverage on our model (observations #6 & 25). We also have one borderline case.  

```{r}
ols_plot_resid_lev(mod3)
```

Once again, county #11 appears to be an outlier with leverage. County #55 is new in terms of outliers. However, we previously flagged County 185 as potentially having data quality issues.  

```{r}
data2 %>% filter(county %in% c(11, 55, 185))
```

We can see that without this observation, we now explain a lot more of the variation (~78% compared to ~72%). Additionally, `taxpc` is no longer significant.  

```{r, warning = F, error = F}
data2_out2 <- data2 %>% filter(county != 11) %>% filter(county != 55) %>% filter(county != 185)

mod3_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2_out2)
stargazer(mod3, mod3_out, type = "text")
```

Although eliminating the outlier has improved our fit and changed the strength of the relationship with `taxpc`, and although counties #11 & 55 have leverage, we do not have any reason to suspect that they have data quality issues. 

What if we only eliminate the county with data quality issues? Now our model fit is slightly less improved (73% compared to the original 72%). The variable `taxpc` is again signficant. 

```{r, warning = F, error = F}
data2_out3 <- data2 %>% filter(county != 185)

mod3_out2 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80, data = data2_out3)
stargazer(mod3, mod3_out, mod3_out2, type = "text")
```

For now, we will leave all of these in our model since #11 & 55 do not appear to have data quality issues and #185 does not change our model signficantly. 


### Model 4

For our fourth model, we aimed to evaluate the robustness of the relationship between crime policy and crime when controlling for many contextual factors. 

We used the following independent variables to predict crime: 

*Crime policy* 

(1) log(`prbarr`)

(2) log(`prbconv`)

*Contextual variables* 

(3) log(`wfed`) 

(4) log(`taxpc`)

(5) `density` - higher density means more opportunity for crime to be committed and reported; denser areas also tend to have more economic activity and so this variable could be related to wages 

(6) `mix` - different types of crime in an area may affect the assertiveness of the criminal justice system; a community may see more arrests and convictions in response to more frightening face-to-face crimes in an area

(7) `pctymle` - young males are more likely to commit crimes and thus areas with higher portions of this type of person may see more crime regardless of their policies 

(8) `pctmin80` - minorities are likelier to be arrested and convicted (and from a lot of research we know that this is not always related to their likelihood to commit a crime)

(9) `west` - Western North Carolina is potentially very different from other regions in North Carolina, as we noted at the beginning of this report. A different culture could relate to different crime patterns and relationships between crime and other factors. For example, this variable is highly related to `pctmin80` (Western counties have lower levels of minorities) and this could affect crime rate through differences in diversity. 

(10) log(`wtrd`) 

(11) log(`prbpris`) 

(12) `central`

(13) `urban`


#### Discussion on results

We can see that our additions did not improve our model fit. 

```{r, warning=F, error=F}
mod4 <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80 + log(wtrd) +  log(`prbpris`) + avgsen + central + urban, data = data2)
stargazer(mod3, mod4, type = "text")
```

#### Outliers

One can see that there is one potential outlier that also has a lot of leverage on our model. 

```{r}
ols_plot_resid_lev(mod4)
```

Once again, county #55 appears to be an outlier with leverage.  

```{r}
data2 %>% filter(county == 55)
```

We can see that without this observation, we now explain slightly more of the variation (~74% compared to ~72%). 

```{r, warning = F, error = F}
data2_out <- data2 %>% filter(county != 55)

mod4_out <- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) +log(taxpc) + density + pctymle + pctmin80 + log(wtrd) +  log(`prbpris`) + avgsen + central + urban, data = data2_out)
stargazer(mod4, mod4_out, type = "text")
```

Although eliminating the outlier has improved our fit, and although this outlier has leverage, we do not have any reason to suspect that it has data quality issues. 

```{r}
ols_plot_resid_lev(mod4_out)
```

## Checking the 6 Classical Linear Model assumptions for our perferred model (Model #3) 

### Linear population model

This assumption is automatically fulfilled because we haven't constrained the error term, i.e we haven't required it to be normal. So there's nothing to check at this point.

## Random Sampling

North Carolina has 100 counties and our dataset has only 90 of them. To check random sampling, we need background knowledge of how why these remaining 10 counties were ommited. Unfortunately, the assignment description does not explain it.

In general, we may be concerned about possible problems with independence. For example, ommited counties are the poorest ones, for which it's more difficult to gather data. Perhaps, if these counties were in the sampe, we could capture better the effect of economic policy variables. 

## No perfect collinearity

There is no need to explicitly check for perfect collinearity, because R will alert us if this rare condition happens.

## Zero Conditional Mean

Let's take a look at the diagnostic plots:

```{r, warning=F, error=F}
plot(mod3, which = 1)
```

Notice that there is only a slight deviation from zero conditional mean, indicated by a not perfecly flat curve. This means that our coefficients maybe be biased. Next section, we discuss some ommited variables candidates and the bias direction they would imply.  

## Homoskedasticity

Our residuals versus fitted values plot seems to indicate heteroskedasticity - thickness for fitted values close to the mean seems to be larger. The scale location plot gives us another way to assess this assumption, and tells a similar story.

```{r, warning=F, error=F}
plot(mod3, which = 3)
```

Given this evidence, we will proceed with robust standard errors. The code below provides a test for statistical significance of our model weights that is robust to heteroskedasticity. We can see that the same pattern we saw previously in model 2 for statistical signficance has continued through model 3. 

* log(`prbarr`)
* log(`prbconv`)
* log(`wfed`)
* `pctmin80` 

The following were not: 

* log(`taxpc`)
* `density`
* `pctymle`

```{r}
coeftest(mod3, vcov. = vcovHC)
```

## Normality of Errors

To check normality of errors, we can look at the qqplot that's part of R's standard diagnostics.

```{r, warning=F, error=F}
plot(mod3, which = 2)
```

We can also visually look at the residuals directly.

```{r, warning=F, error=F}
hist(mod3$residuals, breaks = 20, main = "Residuals from Linear Model Predicting log(Crime Rate)")
```

Both methods suggest we have a leftward skew. However, we have a large sample size, so the CLT tells us that our estimators will have a normal sampling distribution. Our look at the histogram confirms that we aren't in a situation with an extreme skew, so n=90 should be sufficient for the CLT.

### Model 5 

Based on our correction for heteroskedasticity, we generated an additional with only those elements with statistical signficance. Although many of the other elements are important in real life, in an effort towards the greatest parsimony and explanatory power we prefer the following model. 

```{r, error=F, warning=F}
mod5<- lm(log(crmrte) ~ log(prbarr) + log(prbconv) + log(wfed) + pctmin80, data = data2)
stargazer(mod5, type = "text")
```

We can see that even with a test robust to heteroskedasticity, all of our elements are statistically significant. 

```{r}
coeftest(mod5, vcov. = vcovHC)
```

We ran the same assumptions tests as above (as evidence by the plots below). We appear to meet all of our assumptions, especially with the inclusion of robust standard errors. 

```{r}
plot(mod5)
```

## A Regression Table

Of our models, the third appears to be our best for predicting crime. We have included both crime and economic variables in the model, as well as some key contextual variables. 

```{r, warning=F, error=F}
stargazer(mod1, mod2a, mod3, mod4, mod5, type = "text")
```


# A Discussion of Other Possible Omitted Variables

Omitted variables can change the strength of relationships between our independent and dependent variables. As noted above, contrary to our expectation the wages of federal employees was *positively* related to crime. We expected that areas with a higher cost of living (proxied by wages of federal employees) would have less crime. However, it appears that those have more crime. There could be other hidden or unmeasured factors that explain this unexpected direction. 

We lacked many variables related to community-level elements that would be expected to affect crime. Could any of these have affected the quantification of the relationships above? 

**Education levels of the populace** 

Although we may be able to proxy education levels by looking at which region a county is in, we would not have a very good measure of overall education levels. A county can have a university, but still have low education levels overall. One would predict that education is negatively related to crime - the more educated, the less crime one would expect to see. Education levels would be positively correlated to wages and cost of living. It's unclear what direction the relationship would be to the distribution of minorities, or to criminal justice practice. Thus, one would guess that the bias would be towards zero (positive relationship to wages * negative relationship to crime = negative). This means that if anything omitting this variable would have diminished the measure of relationships.

**Public works & investments** 

One would predict that, like education, the relationship between and public works and investments would be negative. More nonprofits and community organizations could strengthen ties in communities, thus lowering crime rates. This factor would probably be inversely related to arrest and conviction rates, since a more punitive society may invest less in rewards to the community. It would probably be positively related to wages, since more money means more funds available to invest back. The direction of the relationship to the distribution of minorities is unclear. Thus, one would guess that the bias would be towards zero (negative relationship to arrest rate * negative relationship to conviction rate * positive relationship to wages * negative relationship to crime = negative).  

**Individual Health (mental and physical)** 

Individual health - both mental and physical - would likely be related to less crime. The healthier you are, the more mental and physical resources you have to solve problems in ways that do not necessitate committing a crime. One would expect this factor to be positively related to wages but negatively related to percent minorities (racism has negative impacts on health). The direction of the relationship to arrest or conviction rates is unclear. The direction of the relationship to density could go either way - people in cities may have access to better and more accessible health care, but live in a more polluted environment. Thus, one would guess that the bias would be away from zero (positive relationship to wages * negative relationship to percent minority * negative relationship to crime = positive). This could have elevated the quantification of our relationships above. Could better individual health be the hidden variable that is influencing our relationship to federal wages (our proxy for cost of living)? 

**Community Health**

Community health is about the cohesiveness of a community. If you have good supports in your community, you would be less likely to commit crime. The relationship to wages is unclear - communities at many income levels can have loving, supportive relationships. It is unclear what the relationship would be to percent minorities. This factor would probably be inversely related to arrest and conviction rates, since a more loving, cohesive community may be less likely to turn to punishment. Thus, one would guess that the bias would be towards zero (negative relationship to arrest rate * negative relationship to conviction rate * negative relationship to crime = negative).

**Inequality** 

We only had measures of average wage by sector in each county. We don't have information about the distribution within and across these sectors. With more inequality, one would expect to see more crime - more anger & therefore reason to commit crime combined with greater levels of deprivation & need to commit crime would naturally lead to more crime. The direction of the relationship between inequality and any of our independent variables in our final model is unclear - thus it is difficult to tell what direction this bias would be in. 

# A Conclusion

Based on our dataset, our analyses suggest that acting on crime deterrants within the criminal justice system (like arrests and convictions) is important policy efforts to take when it comes to preventing crime. Helpful follow-ups in this investigation would be to leverage variation across time for conducting a difference-in-differences test or large changes in policy that would help with regression discontinuity. However, our findings can be used to hone in on future policy changes that can be evaluated with more rigorous causal analyses. 

It would also be helpful to follow-up on this investigation with the addition of health variables. These omitted variables could have affected the relationships tested here. Specifically, areas with higher cost of living may have people living there with better health care and less stress, which would make them healthier and give them less reason to commit crimes. 

Other factors outside of policy - such as the distribution of minorities - are also important when it comes to crime. Communities with higher rates of either may require additional resources or different types of policies to prevent crime. For example, if the actual number of crimes is different from the number of reported crimes -- for example, in a community with a lot of racial conflict where some groups over-report others to the police -- we would see elevated measures of crime. If policies aim to reduce actual crime (rather than just reported crime), we would need more nuanced measures to track these and would need to turn towards more nuanced ways to address this (perhaps through strenghening group cohesiveness and hosting workshops to reduce racism). 

Some policies could include: 

* Improved training and hiring of criminal justice staff, that could improve the ratio of arrest and convictions to crimes. Making sure that those who commit crimes are caught seems to be important in deterring crime - whether by keeping people off the streets, rehabilitating them, or scaring others away from commiting crimes 

* Even though we could not find much support for the importance of economic policy variables to prevent crimes, measures to improve living conditions locally should not detract from efforts on the criminal justice system front. Targeting economic development policies that can bring higher wages and improve the quality of life in a region is likely important to deter crime. We did find that the wages of federal employees appear positively correlated with crime - one would imagine that counties would not intend to enact policies that depress wages in their region. There are likely hidden or unmeasured factors (such as health, inequality, corruption, or high cost of living without investent back into the community) that could be driving this effect. 

* We found that communities with higher density of minorities have higher rates of crime. More follow-up would be needed to investigate mediators of this effect - such as racism, which could lead to higher rates of reported crime (even when holding actual crime constant) - but efforts towards community investments such as those related to the economic development efforts could also be used to fight crime. 

* Both of these streams (criminal and economic reform) could be folded under the umbrella of investing more in one's community. It is not enough to punish and deter - one needs to also get at the root causes of crime and why someone may be compelled to commit a crime in the first place. The fact that crime is higher with high rates of minorities suggests that the picture is more complicated than it may seem when looking at crime policy variables. Are there elements of community cohesion that could be improved to help reduce crime? Or efforts to mitigate the systemic effects of racism? Perhaps more crimes are simply reported in those areas - so the elevate numbers are not truly a reflection of actual crime committed but rather enhanced surveillance. 



